---
runtime: shiny
title: "Choropleth Map by USA Zipcode Level"
output: 
  flexdashboard::flex_dashboard:
  orientation: columns
vertical_layout: scroll
social: [ "twitter", "facebook", "menu" ]
source_code: embed
theme: flatly

---
  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Load packages and initialize data here
library(flexdashboard)
library(rgdal)
library(leaflet)
library(dplyr)
library(magrittr)
library(readr)
library(knitr)
library(shiny)
```

Map
=========================
  
Sidebar {.sidebar} 
----------------------------
Here is a leaflet map showing the estimated population at a state-level by zipcode. The data was obtained from [USA Zipcode](http://www.unitedstateszipcodes.org). 

```{r, echo=FALSE}
hr()

selectInput("state", label = "Select a state: ",
            choices = state.abb, selected = "NY")

```

Made for fun by [Jasmine Dumas](http://jasdumas.github.io/)

Map {data-height=1000}
----------------------------
  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# USA level zipcode for 2015
load("usa.RData")

# read in csv file with state specific state data
all_usa_zip <- read_csv("zip_code_database.csv")


# state specific zipcodes and data
STATE_SHP <- reactive({
  
  
  # filter down to the specific state and select 
  
  specific_state <- all_usa_zip %>% dplyr::filter(state == input$state) %>% 
    select(zip, primary_city, county, irs_estimated_population_2014)
  
  # change the column names for the merge by arguement
  
  colnames(specific_state) <- c("zipcode", "City", "County", "irs_estimated_population_2014")
  
  # change the zipcode to a factor to full join the table
  
  specific_state$zipcode = factor(specific_state$zipcode)
  
  # remove redundant County label in County column
  
  specific_state$County <- gsub("County", "", specific_state$County) 
  # there is probably a space after now but not concerned about that
  
  # full join the data set 
  
  state_join <- dplyr::full_join(usa@data, specific_state)
  
  # remove rows with NA's - i.e. remove everything except the choosen state
  
  state_clean = na.omit(state_join)
  
  # Merge a Spatial object having a data.frame (i.e. merging of non-spatial attributes).
  # all.x = F removes NA values that are not common to both datasets
  
  STATE_SHP <- sp::merge(x=usa, y=state_clean, all.x = F)
  
}) # END OF REACTIVE

# color palette
renderLeaflet({
  
  #   pal <- colorNumeric(
  #   palette = "Blues",
  #   domain = STATE_SHP()$irs_estimated_population_2014
  # )
  pal <- colorBin(palette = "BuPu", domain = STATE_SHP()$irs_estimated_population_2014, bins = 8)
  
  state_popup <- paste0("County: ", 
                        STATE_SHP()$County, 
                        "City: ", 
                        STATE_SHP()$City, 
                        "Est. Population: ",
                        STATE_SHP()$irs_estimated_population_2014)  
  # plot the map
  leaflet(data = STATE_SHP()) %>%
    addProviderTiles("CartoDB.Positron") %>%
    addPolygons(fillColor = ~pal(irs_estimated_population_2014), 
                fillOpacity = 0.7, 
                color = "#BDBDC3", 
                weight = 1, 
                popup = state_popup) %>%
    addLegend("bottomleft", 
              pal = pal, 
              values = ~irs_estimated_population_2014,
              title = "Est. Population",
              opacity = 1)
  
  
})
